---
title: "Stats 141XP"
author: "Brian"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(tidyverse)
```


```{r}
df <- read.csv("wahl_data.csv")
```

# Correlation of courses taken against classes

```{r}
df_selected <- df %>%
  select(count_of_courses_taken, arts_ct, engineering_ct, humanities_ugeduc_ct,
         life_sci_ct, phys_sci_ct, soc_sci_intl_inst_ct, mgmt_ct, other_prof_ct)

cor_matrix <- cor(df_selected, use = "complete.obs")

cor_matrix["count_of_courses_taken", ]

```

```{r}
cor_df <- as.data.frame(cor_matrix["count_of_courses_taken", ])
cor_df$subject <- rownames(cor_df)
colnames(cor_df)[1] <- "correlation"

cor_df <- cor_df %>% filter(subject != "count_of_courses_taken")

ggplot(cor_df, aes(x = reorder(subject, correlation), y = correlation)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Correlation with Total Courses Taken",
       x = "Subject Area",
       y = "Correlation") +
  theme_minimal()

```
Arts have a moderately positive correlation with Total Courses taken with UCLA


# Average courses per subject, separated by online or not

```{r}
df %>%
  group_by(all_online) %>%
  summarise(across(c(arts_ct, engineering_ct, humanities_ugeduc_ct, 
                     life_sci_ct, phys_sci_ct, soc_sci_intl_inst_ct, 
                     mgmt_ct, other_prof_ct), mean, na.rm = TRUE))
```


```{r}
# Pivot for plotting
df_long_online <- df %>%
  group_by(all_online) %>%
  summarise(across(ends_with("_ct") & !c(online_course_ct, lower_division_course_ct, count_of_courses_taken), 
                   mean, na.rm = TRUE)) %>%
  pivot_longer(-all_online, names_to = "subject", values_to = "mean_courses")

# Plot
ggplot(df_long_online, aes(x = subject, y = mean_courses, fill = factor(all_online))) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(title = "Average Courses per Subject: Fully Online vs Not",
       x = "Subject", y = "Average Courses",
       fill = "Fully Online (1 = Yes)") +
  theme_minimal()
```

Arts, social sciences, management, humanities, life sciences are more popular online
Physical sciences and engineering are more popular offline

# Lower Division Courses percentage

```{r}
df_lower <- df %>%
  mutate(pct_lower_div = lower_division_course_ct / count_of_courses_taken)

summary(df_lower$pct_lower_div)
```
```{r}
ggplot(df_lower, aes(x = pct_lower_div)) +
  geom_histogram(binwidth = 0.1, fill = "steelblue", color = "white") +
  labs(title = "Distribution of Lower Division Course Proportion",
       x = "Proportion of Lower Division Courses",
       y = "Number of Students") +
  theme_minimal()
```

# Arts vs Non Arts Lower Div

```{r}
df_arts <- df %>% 
   filter(arts_ct %in% c(2, 4))

df_arts <- df_arts %>%
  mutate(pct_lower_div = lower_division_course_ct / count_of_courses_taken)

df_non_arts <- df %>% 
  filter(arts_ct %in% c(0, 1)) %>%
  mutate(pct_lower_div = lower_division_course_ct / count_of_courses_taken)

# Compare mean proportions
mean(df_arts$pct_lower_div, na.rm = TRUE)
mean(df_non_arts$pct_lower_div, na.rm = TRUE)

t.test(df_arts$pct_lower_div, df_non_arts$pct_lower_div)
```

```{r}
df %>%
  mutate(pct_lower_div = lower_division_course_ct / count_of_courses_taken,
         group = ifelse(arts_ct %in% c(2, 4), "Took Arts", "No Arts")) %>%
  ggplot(aes(x = group, y = pct_lower_div, fill = group)) +
  geom_boxplot() +
  labs(title = "Proportion of Lower Division Courses: Arts vs Non-Arts",
       y = "Proportion Lower Division") +
  theme_minimal()
```


